---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# infomax

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)

[![R-CMD-check](https://github.com/craddm/infomax/workflows/R-CMD-check/badge.svg)](https://github.com/craddm/infomax/actions)
[![codecov](https://codecov.io/gh/craddm/infomax/branch/master/graph/badge.svg?token=8VYL66NE7Z)](https://codecov.io/gh/craddm/infomax)
[![R-CMD-check](https://github.com/eegverse/infomax/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/eegverse/infomax/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The `infomax` package is an R implementation of the Infomax and Extended Infomax algorithms (Bell & Sejnowski, 1995; Makeig, Bell, Jung, & Sejnowski, 1996) for Independent Component Analysis. 

## Installation

The package is not currently released on [CRAN](https://CRAN.R-project.org).

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("craddm/infomax")
```

## Example

```{r example}
library(infomax)
## basic example code
```

ICA can be used to separate linear mixtures of different, independent signal sources.

In the example below, we generate two independent signals and then mix them together.

```{r make-example-objects}
time_x <- seq(0, 1, by = 1/256)
source_a <- sin(2 * pi * 5 * time_x)
source_b <- sin(2 * pi * 10 * time_x)
plot(time_x, source_a, type = "l")
plot(time_x, source_b, type = "l")
plot(time_x, source_a + 2 * source_b, type = "l")
plot(time_x, source_a * 3.4 + 1.5 * source_b, type = "l")
```

The function returns the estimated mixing matrix, unmixing matrix, and unmixed source timecourses.

```{r example-figures}
mixed_data <- matrix(NA,
                     nrow = length(time_x),
                     ncol = 2)
mixed_data[, 1] <- source_a + 2 * source_b
mixed_data[, 2] <- source_a * 3.4 + 1.5 * source_b
dat_out <- run_infomax(mixed_data, whiten = "PCA")
plot(time_x,
     dat_out$S[,1],
     type = "l")
plot(time_x,
     dat_out$S[,2],
     type = "l")
```

## References

* Bell, A.J., & Sejnowski, T.J. (1995). An information-maximization approach to blind separation and blind deconvolution. *Neural Computation, 7,* 1129-159
* Makeig, S., Bell, A.J., Jung, T-P and Sejnowski, T.J., "Independent component analysis of  electroencephalographic data,"  In: D. Touretzky, M. Mozer and M. Hasselmo (Eds). Advances in Neural  Information Processing Systems 8:145-151, MIT Press, Cambridge, MA (1996).
